---
title: "Market Basket Analysis"
author: "Rohan Stanly Keecheril"
format: html
editor: visual
---

# Market Basket Analysis

## 1. Introduction

Market basket analysis is a method used to understand purchasing behavior/ choices made by customers. Based on frequency of purchases (Support) and the strength of association between items (Confidence), rules can be developed to predict which items are likely to be found together in a customer’s basket.

## 2. Example of Market Basket Analysis using groceries data set from Kaggle.

### 2.1 Data Exploration

```{r echo=FALSE, message=FALSE, warning=FALSE}
Groceries_data<-read.csv("C:/Users/rstan/Desktop/MBA/T1/Introduction to Business Analytics/R docs/Groceries_dataset.csv")
head(Groceries_data)
tail(Groceries_data)
str(Groceries_data) #To see structure of data
length(unique(Groceries_data$Member_number))
```

The Groceries_data dataset from Kaggle contains 38,765 observations and 3 features: Member_number, Date of purchase, and ItemDescription.

### 2.2 Association Rule Mining

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr) #Call dplayer
transaction_data<-Groceries_data%>%
  group_by(Member_number,Date)%>%
  summarize(items=list(itemDescription),.groups="drop")
library(arules)
#transaction_data
#create transaction object
transactions_list<-as(transaction_data$items,"transactions")
#create association rules
rules<-apriori(transactions_list,parameter=list(support=0.0005,conf=0.2,minlen=2))
inspect(head(rules,10))
```

Association rules were generated with a minimum support of 0.0005, a minimum confidence of 0.2, and a minimum length of 2. The results indicate that approximately 10 items are associated with one another, with the majority of these associations involving whole milk.

### 2.3 Results

Using association rule mining with support = 0.0005, confidence = 0.2, and minimum length = 2, about ten rules were identified. Most of these rules involve whole milk, which frequently appears as a common item in customer baskets. An exception was the association between spices and soda, suggesting occasional niche purchasing patterns.

### 2.4 Inference

The strong presence of whole milk highlights its role as a staple product and a driver of other purchases. Retailers can use this insight for product bundling, cross-selling, or store placement strategies. The spices–soda rule, while less frequent, indicates opportunities to target special consumption occasions.

### 2.5 Conclusions

Market basket analysis reveals that whole milk dominates customer baskets, making it a key product for promotions and cross-selling. At the same time, smaller associations point to niche behaviors that businesses can leverage for targeted marketing.
